A Dissection of Foregrounds in Epoch of Reionization Observations from the Murchison Widefield Array

### 2015-01-19 ms.tex
Addressed most of the referee comments and made lots of minor changes throughout

Detection of 21~cm emission of neutral hydrogen from the epoch of reionization, at redshifts $z>6$, is limited primarily by foreground emission.  We investigate the chromatic signatures of an all--sky foreground model imprinted by an instrumental transfer function. Using the delay spectrum technique {\bf that maps the Fourier transform of the measurements to locations of foreground objects on the celestial sphere through delays in signals received by antenna pairs}, we demonstrate that the foreground signatures that have the largest impact on the H{\sc i} signal arise from power received far away from the primary field of view. Comparing data from recent Murchison Widefield Array observations with simulations separated into components based on type of emission, we identify diffuse emission near the horizon as a significant contributing factor, even on wide antenna spacings {\bf which are usually not sensitive to diffuse emission}. For signals entering through the primary field of view, compact emission dominates the foreground contamination. These two mechanisms imprint a characteristic {\it pitchfork} signature on the ``foreground wedge'' in Fourier space --- {\bf a signature generic to all wide--field measurements}. Based on these results, we propose that selective down--weighting of data based on antenna spacing vector and time will mitigate foreground contamination substantially (up to a {\bf factor $\sim 100$} with negligible loss of sensitivity) in redshifted 21~cm power spectrum analysis. {\bf A dish--shaped antenna yields the most desirable features in terms of foreground contamination, relative to a dipole or a phased array}.

@@

Considerable effort has been and continues to be made towards understanding the $k$--space behavior of foreground signatures in the observed power spectrum and formulating robust estimators of the true power spectrum \citep{bow09,liu09,dat10,liu11,mor12,tro12,pob13,thy13,dil14,liu14a,liu14b}. {\bf A model that provides a generic explanation for the observed foreground power spectrum has emerged, whereby the wide--field (and chromatic) response of the instrument causes the power in smooth spectrum foregrounds to occupy higher $k$--modes into the so-called `wedge'}. % A general model of scattering of smooth spectrum foregrounds to higher $k$--modes by the wide--field (and chromatic) response of the instrument, into the so-called `wedge', has emerged that provides a high level explanation for the observed foreground power spectrum. 
The conservative foreground strategy that has developed alongside this work is to discard $k$--modes which could be contaminated ({\it avoidance}). The more aggressive alternative is to subtract a sky model and regain access to modes discarded by {\it avoidance}. In both cases, which parts of the sky are most critical to either {\it avoid} or {\it subtract} has remained largely uncertain. Here, we focus primarily on extending the {\it avoidance} strategy by identifying foreground components at greatest risk to ``leak'' from foreground modes to EoR modes and proposing a scheme for down--weighting these components.

@@

Interferometer array data known as {\it visibilities}, $V_{bf}\equiv V_{bf}(\boldsymbol{b},f)$, represent correlations between time--series of electric fields measured by different antenna pairs with separation vectors $\boldsymbol{b}$ and then Fourier transformed along the time axis to obtain a spectrum along the frequency ($f$) axis. If $I(\hat{\boldsymbol{s}},f)$ and $A(\hat{\boldsymbol{s}},f)$ are the sky brightness and antenna's directional power--pattern, respectively, at different frequencies as a function of direction on the sky denoted by the unit vector ($\hat{\boldsymbol{s}}$), and $W_f\equiv W_f(f)$ denotes instrumental bandpass weights, then $V_{bf}$ can be written as (with a slight adaptation from \citet{van34}, \citet{zer38}, and \citet{tho01}):
\begin{align}\label{eqn:obsvis}
  V_{bf} &= \iint\limits_\textrm{sky} A(\hat{\boldsymbol{s}},f)\,I(\hat{\boldsymbol{s}},f)\,W_f\,e^{-i2\pi f\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}}\,\dif\Omega,
\end{align}
where, $c$ is the speed of light, and $\dif\Omega$ is the solid angle element to which $\hat{\boldsymbol{s}}$ is the unit normal vector. We wish to {\bf note} that this equation is valid {\bf even for wide--field applications where the sky cannot be approximated to be flat}.

@@

Without loss of generality, the phase center can be assumed to be the zenith relative to the local tangent plane. Then $\boldsymbol{u}$ lies on this plane for measurements constrained to be on it. If the array of antennas are also coplanar lying on the local tangent plane, then $\boldsymbol{u}=\boldsymbol{b}/\lambda$. Under such circumstances, equations~\ref{eqn:obsvis} and \ref{eqn:delay} closely resemble equations~\ref{eqn:vis} and \ref{eqn:los-transform} respectively. However, they are not quite identical to each other. The difference basically arises from the fact noted above that $\boldsymbol{b}$ is independent of frequency while $\boldsymbol{u}$ is not. \citet{par12b} and \citet{liu14a} have discussed the mathematical correspondence between the two.

% Radio emission from foreground objects, due to their spatial and spectral (chromatic) structure, contaminate the received signal \citep{bow09,dat10,mor12,tro12,thy13,liu14a,liu14b}. 
Foregrounds can be described in both the delay spectrum and Fourier frameworks. For our study, which primarily concerns with foreground characterization, we find the former to be simple and yet extremely useful, while maintaining a close correspondence with the latter despite subtle differences. In order to express a quantity derived from $V_{u\eta}$ that has units identical to that of aforementioned cosmological H{\sc i} power spectrum, we define:
\begin{align}\label{eqn:fake-power-spectrum}
  P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel) &\equiv |V_{b\tau}|^2\left(\frac{A_\textrm{e}}{\lambda^2\Delta B}\right)\left(\frac{D^2\Delta D}{\Delta B}\right)\left(\frac{\lambda^2}{2k_\textrm{B}}\right)^2,
\end{align}
where, $\boldsymbol{u}$ and $\eta$ in equations~\ref{eqn:k-perp} and \ref{eqn:k-prll} have been correspondingly replaced with $\boldsymbol{b}/\lambda$ and $\tau$ respectively.

In summary, the delay spectrum, $V_{b\tau}$, is obtained from {\it visibilities} which are the basic data blocks measured by each antenna pair, using equations~\ref{eqn:obsvis} and \ref{eqn:delay-transform}. $V_{b\tau}$ captures all the effects of EoR H{\sc i} signal corruption caused by foregrounds and the instrument. At the same time, it is closely related to the sought power spectrum $P(\boldsymbol{k}_\perp,k_\parallel)$ containing critical information about spatial scales. In this paper, we often use the quantity $P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel)$, and sometimes $V_{b\tau}$.

@@

\subsection{Observation Parameters}\label{sec:obsparms}

@@

The data are flagged\footnote{\url{http://sourceforge.net/p/aoflagger}} for interference \citep{off10,off12}, removing 3\% of the data and averaged in time  and frequency from the raw 0.5~s, 40~kHz to 2s, 80~kHz. These data are then calibrated to a simulation of the sky containing 2420 point--like objects selected from the MWA Commissioning Survey \citep[MWACS;][]{hur14}. It has a flux density limit of 25~mJy and a declination range of $-12\arcdeg$~ to $-40\arcdeg$~ evenly covering the field of view of the observations reported here. The objects used in calibration are selected to lie inside the 5\% contour of the primary lobe of the tile power--pattern. The calibration algorithm -- based on forward modeling software by \citet{sul12} and the calibration method described by \citet{sal14} -- computes complex gain solutions per channel per antenna averaged to two minute intervals. The solutions are fairly low signal--to--noise so we iteratively average along the antenna and frequency dimensions to capture the relatively independent passband and antenna--to--antenna variation. First, we average the channel gains over all antennas to obtain a high signal--to--noise measurement of the bandpass. After applying this single passband, we do a second round of calibration and fit second and first order polynomials for amplitude and phase respectively for each antenna. This flattens any residual variation in bandpass and removes small phase slopes due to variations in cable delay. Finally, we fit for an additional phase known to be caused by small reflections in a subset of cables.

@@

We show {\bf delay power spectra} from MWA data after deconvolution along the delay axis in Figure~\ref{fig:fhd_data}. The principle feature is the {\it foreground wedge}, whose boundaries, marked by white lines, increase linearly with light travel time corresponding to baseline length. Note that the wedge is not filled out uniformly. Each baseline has a different orientation and resulting response to different objects in the sky. It is this variation that we will explore below.

The {\it off--zenith} pointing {\bf has notably higher power in the {\it foreground wedge} than in the {\it zenith} pointing}. This is shown later in \S\ref{sec:diffuse} to be due to response of the array to the bright Galactic center and Galactic plane in the westward sky. Consequently, the contamination into the {\it EoR window} is also higher, especially on short antenna spacings. Emission in the {\it zenith} pointing is more centrally concentrated {\bf due to a zenith--centered power--pattern and faster suppression of response away from zenith, whereas the {\it off--zenith} pointing has a power--pattern phased eastward of zenith and is also more responsive to emission farther away from the primary lobe relative to that in the {\it zenith} pointing. The power pattern is discussed in detail in \S\ref{sec:instrument_model} and Figure~\ref{fig:power_pattern}}.

@@

The instrument model consists primarily of the tile beam power--pattern, $A(\hat{\boldsymbol{s}},f)$ (see equation~\ref{eqn:obsvis}). It is modeled as a mutually--coupled 4--by--4 dipole array with the overall power--pattern of each individual dipole calculated via finite element electromagnetic simulation \citep{sut14_arXiv}. To speed up simulations, we find that a phased array of isotropic radiators at a height of 0.3~m above an infinite ground plane provides a very good approximation to the full simulation, hence we use the idealized dipoles. We also assume that each individual dipole signal has random delay fluctuations of rms 0.05~ns, a number in line with the known repeatability and stability level of the analog signal chain \citep{bow07b}. Besides having the effect of adding a time--dependent uncertainty in the power pattern, these random delay fluctuations reduce the coherence in the phased addition of dipole signals resulting in deviations from predicted models of the power--pattern, most prominently at its nulls. The power--patterns so obtained are shown in Figure~\ref{fig:power_pattern} for the pointings chosen in our study.

\begin{figure}[htb]
\centering
% \includegraphics[width=\linewidth]{figures/v1_0/delta_array_powerpattern_0.3m_ground_custom.eps}
% \includegraphics[width=\linewidth]{f4.eps}
\includegraphics[width=\linewidth]{fig4.eps}
\caption{Models of MWA tile power--pattern for {\it off--zenith} (top) and {\it zenith} pointings (bottom) at 185~MHz. An MWA tile is modeled as a 4$\times$4 array of isotropic radiators placed 0.3~m above an infinite ground plane. Random fluctuations of rms 0.05~ns have been added to delay corrections during the phased addition of voltages from these isotropic radiators. The logarithmic color scale (shown on the right) is common to both panels. \label{fig:power_pattern}}
\end{figure}

@@

A low resolution version of the diffuse foreground model used is shown in Figure~\ref{fig:DSM}. Contours of the MWA tile power--pattern shown in Figure~\ref{fig:power_pattern} are also overlaid. Of notable significance in the {\it off--zenith} pointing is the presence of a portion of the Galactic plane and the bright Galactic center in the westward sky, where the MWA tile power gain is significant ($\gtrsim 12$\%). In the {\it zenith} pointing, the Galactic plane has set and the power--pattern in that direction is at least 16 times smaller.

\begin{figure}[htb]
\centering
% \includegraphics[width=\linewidth]{figures/v1_0/dsm.eps}
% \includegraphics[width=\linewidth]{f5.eps}
\includegraphics[width=\linewidth]{fig5.eps}
\caption{Sky brightness temperature of the diffuse foreground model at 185~MHz visible during {\it off--zenith} (top) and {\it zenith} (bottom) pointings. The color scale on the right is logarithmic and is common to both panels. Contours of power--pattern shown in Figure~\ref{fig:power_pattern} are overlaid. The contour levels shown are 0.00195, 0.00781, 0.0312, 0.125, and 0.5. The Galactic center and a portion of the Galactic plane are prominently visible during the {\it off--zenith} pointing and the MWA tile power gain is significant ($\gtrsim 12$\%) in that direction. In contrast, emission from the Galactic plane in {\it zenith} pointing is significantly lesser. \label{fig:DSM}}
\end{figure}

@@

\begin{figure}[htb]
\centering
% \includegraphics[width=\linewidth]{figures/v1_0/csm.eps}
% \includegraphics[width=\linewidth]{f6.eps}
\includegraphics[width=\linewidth]{fig6.eps}
\caption{Flux densities of bright point sources at 185~MHz visible during {\it off--zenith} (top) and {\it zenith} (bottom) pointings. The color scale on the right is logarithmic and is common to both panels and corresponds to the color--coded filled circles. Contours of power--pattern shown in Figure~\ref{fig:power_pattern} are overlaid. The contour levels shown are 0.001953125, 0.0078125, 0.03125, 0.125, and 0.5. \label{fig:CSM}}
\end{figure}

@@

In addition to intrinsic model uncertainty, delay spectra from simulations and data each have fluctuations due to thermal noise in the delay spectrum with rms $\sim$ 1.4$\times 10^7$~Jy~Hz. We estimate the ratio of delay spectra from data and simulations as $\rho = |V^\textrm{D}_{b\tau}(\boldsymbol{b},\tau)|\,/\,|V^\textrm{S}_{b\tau}(\boldsymbol{b},\tau)|$, where superscripts D and S denote data and simulation, respectively.

The median absolute deviation of $\log_{10}\rho$ inside the {\it foreground wedge} for both pointings is $\approx 0.28$. This corresponds to $\sim$~90\% fractional difference between data and modeling on average with either pointing. We also simulated delay spectra after assigning spectral indices drawn randomly from a {\it gaussian} distribution with a mean of $\langle\alpha_\textrm{sp}\rangle=-0.83$ and a HWHM of $\Delta\alpha_\textrm{sp}=0.35$ to the point sources in our compact foreground model. These simulations typically yielded a median absolute deviation of $\approx 0.29$ for $\log_{10}\rho$ indicating fractional differences of $\sim$~95\%. This demonstrates that a fractional deviation of $\sim$~90\% observed between data and simulations is in line with expectations when the aforementioned uncertainty in foreground models, thermal noise fluctuations in measurements, and uncertainties in antenna power pattern due to random delay fluctuations are taken into account. Currently, we are significantly limited by the unavailability of all--sky foreground models at the observing frequency of 185~MHz.

These uncertainties are presented only to confirm the qualitative agreement already seen between Figures~\ref{fig:fhd_data} and \ref{fig:sim_data}. These estimates are conservative. A full treatment of all uncertainties and deviations from ideal behavior such as frequency dependent errors in tile power--pattern \citep{ber14}, calibration \citep{dat10}, data corruption due to interference, anisoplanatic wide--field imaging and ionospheric effects \citep{int09} will bring the simulations much closer in agreement with observations, but is beyond the scope of this paper. {\bf Hereafter}, our primary objective is to explore in detail the foreground signatures embedded in the {\it foreground wedge}.

@@

\item {\it Baseline Orientation}: Since the spatial structure of our foreground model is not expected to be isotropic, we divide our antenna spacings by their orientation, $\theta_\textrm{b}$. We use the {\bf four following bins sweeping 45\arcdeg each}: $-67.5\arcdeg\le \theta_\textrm{b} < -22.5\arcdeg$, $-22.5\arcdeg\le \theta_\textrm{b} < 22.5\arcdeg$, $22.5\arcdeg\le \theta_\textrm{b} < 67.5\arcdeg$, and $67.5\arcdeg\le \theta_\textrm{b} < 112.5\arcdeg$. The bin centers are oriented towards South--East, East, North--East, and North respectively. {\bf While the bin boundaries are arbitrary, they were so chosen to allow bin centers to be aligned along the cardinal and ordinal directions in one half--plane}. Since delays depend on $\theta_\textrm{b}$, binning by $\theta_\textrm{b}$ allows us to match {\bf features of the delay power spectra} to different sky directions.

\item {\it Tile Pointing and Power--Pattern}: Since the MWA tiles are steered electronically, the power--pattern of the tile changes in any observing mode that tracks the {\bf pointing center}. In our study, we take into account the effect of the changing tile power--pattern on the {\bf delay power spectra} (Figure~\ref{fig:power_pattern}). The LST of our data ranges from $\sim$~21 hours through $\sim$~2 hours, and for our analysis we chose the {\it off--zenith} and {\it zenith} pointings shown in Figure~\ref{fig:pointings}. We will show the secondary lobe structures in the tile power--pattern are tied to the foreground signatures in Fourier space.

@@

In subsequent sections, we provide a detailed explanation of our results as a combination of factors noted above. Note that numerous features may overlap at different degrees of significance depending on combinations of parameters. We assign the features to their predominant causes. Secondly, we have used noiseless cases to clearly illustrate the observed foreground signatures. With the addition of noise in the visibilities, some of the weaker features may not be as prominently visible. Since the foreground signatures are far too numerous and they reside in a vast volume of parameter space such as antenna spacing length and orientation, power--pattern, patch of sky under observation, and instrumental configuration, we highlight only some examples of the most notable features observed in the delay spectra of foreground emission.

@@

The most prominent signature seen in the {\it off--zenith} pointing (top panel of Figure~\ref{fig:noiseless-dsm-delay-spectrum}) is due to the bright Galactic center situated on the western horizon co--located with one of the bright secondary lobes of the power--pattern. It appears as a bright branch in the {\bf delay power spectrum} near the negative delay horizon delay limit. This feature is strongest at short antenna spacings and fades with increasing antenna spacing. The {\bf bright} signature is absent in the {\it zenith} pointing (bottom panel, Figure~\ref{fig:noiseless-dsm-delay-spectrum}) because the Galactic center is below the horizon.

@@

The amplitude response of an interferometer to a point source is, to first order, flat with respect to baseline length. Since the main lobe of the power--pattern in the {\it off--zenith} pointing is centered eastward of zenith, the bulk of the compact foreground emission is seen in a branch with positive delay corresponding to the position of the primary lobe of the power--pattern. In the {\it zenith} pointing, compact emission from the same patch of sky is seen as a bright horizontal arm at zero delay  since the primary lobe of the power--pattern is centered at zenith.

>>>>

Detection of 21~cm emission of H{\sc i} from the epoch of reionization, at redshifts $z>6$, is limited primarily by foreground emission.  We investigate the chromatic signatures of an all--sky foreground model imprinted by an instrumental transfer function. Using the delay spectrum technique {\bf that maps the measurements to foreground object locations through signal delays between antenna pairs}, we demonstrate that the foreground signatures that have the largest impact on the H{\sc i} signal arise from power received far away from the primary field of view. Comparing data from recent Murchison Widefield Array observations with simulations {\bf broken down by types of emission}, we identify diffuse emission near the horizon as a significant contributing factor, even on wide antenna spacings {\bf which are usually not sensitive to diffuse emission}. For signals entering through the primary field of view, compact emission dominates the foreground contamination. These two mechanisms imprint a characteristic {\it pitchfork} signature on the ``foreground wedge'' in Fourier space, {\bf which is found to be generic to wide--field measurements}. Based on these results, we propose that selective down--weighting of data based on antenna spacing and time {\bf can} mitigate foreground contamination substantially (up to a {\bf factor $\sim 100$} with negligible loss of sensitivity). {\bf A dish--shaped antenna yields the most desirable features from a foreground contamination viewpoint, relative to a dipole or a phased array}.

@@

Considerable effort has been and continues to be made towards understanding the $k$--space behavior of foreground signatures in the observed power spectrum and formulating robust estimators of the true power spectrum \citep{bow09,liu09,dat10,liu11,mor12,tro12,pob13,thy13,dil14,liu14a,liu14b}. {\bf A model that provides a generic explanation for the observed foreground power spectrum has emerged, whereby the wide--field (and chromatic) response of the instrument causes the power in smooth spectrum foregrounds to occupy higher $k$--modes into the so-called `wedge'}. % A general model of scattering of smooth spectrum foregrounds to higher $k$--modes by the wide--field (and chromatic) response of the instrument, into the so-called `wedge', has emerged that provides a high level explanation for the observed foreground power spectrum. 
The conservative foreground strategy, referred to as {\it avoidance}, that has developed alongside this work is to discard $k$--modes which could be contaminated \citep[for e.g.,][]{par14}. The more aggressive alternative is to subtract a sky model and regain access to modes discarded by {\it avoidance}. In both cases, which parts of the sky are most critical to either {\it avoid} or {\it subtract} has remained largely uncertain. Here, we focus primarily on extending the {\it avoidance} strategy by identifying foreground components at greatest risk to ``leak'' from foreground modes to EoR modes and proposing a scheme for down--weighting these components.

@@

Interferometer array data known as {\it visibilities}, $V_{bf}\equiv V_{bf}(\boldsymbol{b},f)$, represent correlations between time--series of electric fields measured by different antenna pairs with separation vectors $\boldsymbol{b}$ and then Fourier transformed along the time axis to obtain a spectrum along the frequency ($f$) axis. If $I(\hat{\boldsymbol{s}},f)$ and $A(\hat{\boldsymbol{s}},f)$ are the sky brightness and antenna's directional power pattern, respectively, at different frequencies as a function of direction on the sky denoted by the unit vector ($\hat{\boldsymbol{s}}$), and $W_f\equiv W_f(f)$ denotes instrumental bandpass weights, then $V_{bf}$ can be written as (with a slight adaptation from \citet{van34}, \citet{zer38}, and \citet{tho01}):
\begin{align}\label{eqn:obsvis}
  V_{bf} &= \iint\limits_\textrm{sky} A(\hat{\boldsymbol{s}},f)\,I(\hat{\boldsymbol{s}},f)\,W_f\,e^{-i2\pi f\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}}\,\dif\Omega,
\end{align}
where, $c$ is the speed of light, and $\dif\Omega$ is the solid angle element to which $\hat{\boldsymbol{s}}$ is the unit normal vector. We wish to {\bf note} that this equation is valid {\bf even for wide--field measurements where the sky cannot be approximated to be flat}.

@@

Without loss of generality, the phase center can be assumed to be the zenith relative to the local tangent plane. Then $\boldsymbol{u}$ lies on this plane for measurements constrained to be on it. If the array of antennas are also coplanar lying on the local tangent plane, then $\boldsymbol{u}=\boldsymbol{b}/\lambda$. Under such circumstances, equations~\ref{eqn:obsvis} and \ref{eqn:delay} closely resemble equations~\ref{eqn:vis} and \ref{eqn:los-transform} respectively. However, they are not quite identical to each other. The difference is because $\boldsymbol{b}$ is independent of frequency while $\boldsymbol{u}$ is not. \citet{par12b} and \citet{liu14a} have discussed the mathematical correspondence between the two.

% Radio emission from foreground objects, due to their spatial and spectral (chromatic) structure, contaminate the received signal \citep{bow09,dat10,mor12,tro12,thy13,liu14a,liu14b}. 
Foregrounds can be described in both the delay spectrum and Fourier frameworks. For our study, which primarily concerns with foreground characterization, we find the former to be simple and yet extremely useful, while maintaining a close correspondence with the latter despite subtle differences. In order to express a quantity derived from $V_{u\eta}$ that has units identical to that of the cosmological H{\sc i} power spectrum, we define delay power spectrum:
\begin{align}\label{eqn:fake-power-spectrum}
  P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel) &\equiv |V_{b\tau}|^2\left(\frac{A_\textrm{e}}{\lambda^2\Delta B}\right)\left(\frac{D^2\Delta D}{\Delta B}\right)\left(\frac{\lambda^2}{2k_\textrm{B}}\right)^2,
\end{align}
where, $\boldsymbol{u}$ and $\eta$ in equations~\ref{eqn:k-perp} and \ref{eqn:k-prll} have been correspondingly replaced with $\boldsymbol{b}/\lambda$ and $\tau$ respectively.

In summary, the delay spectrum, $V_{b\tau}$, is obtained from {\it visibilities} which are the basic data blocks measured by each antenna pair, using equations~\ref{eqn:obsvis} and \ref{eqn:delay-transform}. $V_{b\tau}$ captures all the effects of EoR H{\sc i} signal corruption caused by foregrounds and the instrument. At the same time, it is closely related to the sought power spectrum $P(\boldsymbol{k}_\perp,k_\parallel)$ containing critical information about spatial scales. % In this paper, we often use the quantity $P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel)$, and sometimes $V_{b\tau}$.

@@

\subsection{EoR Observations}\label{sec:obsparms}

{\bf The MWA is expected to be sensitive to the power spectrum of the H{\sc i} signal over the redshift range $6<z<10$ \citep{thy13,bea13}. Over 600~hours have been currently observed using the MWA with the objective of characterizing the nature of the sources that are responsible for ionizing the IGM, charting the evolution of the global neutral fraction, and probing the nature of quasar emissions by constraining the properties of their ionized proximity zones.}

@@

The data are flagged\footnote{\url{http://sourceforge.net/p/aoflagger}} for interference \citep{off10,off12}, removing 3\% of the data and averaged in time  and frequency from the raw 0.5~s, 40~kHz to 2s, 80~kHz. These data are then calibrated to a simulation of the sky containing 2420 point--like objects selected from the MWA Commissioning Survey \citep[MWACS;][]{hur14}. It has a flux density limit of 25~mJy and a declination range of $-12\arcdeg$~ to $-40\arcdeg$~ evenly covering the field of view of the observations reported here. The objects used in calibration are selected to lie inside the 5\% contour of the primary lobe of the tile power pattern. The calibration algorithm -- based on forward modeling software by \citet{sul12} and the calibration method described by \citet{sal14} -- computes complex gain solutions per channel per antenna averaged to two minute intervals. The solutions are fairly low signal--to--noise so we iteratively average along the antenna and frequency dimensions to capture the relatively independent passband and antenna--to--antenna variation. First, we average the channel gains over all antennas to obtain a high signal--to--noise measurement of the bandpass. After applying this single passband, we do a second round of calibration and fit second and first order polynomials for amplitude and phase respectively for each antenna. This flattens any residual variation in bandpass and removes small phase slopes due to variations in cable delay. Finally, we fit for an additional phase known to be caused by small reflections in a subset of cables.

@@

Figure~\ref{fig:fhd_data} shows {\bf delay power spectra} from MWA data after deconvolution along the delay axis. The principal feature is the {\it foreground wedge}, whose boundaries, marked by white lines, increase linearly with light travel time proportional to baseline length. Note that the wedge is not filled out uniformly. Each baseline has a different orientation that determines the response to different objects in the sky. It is this variation that we will explore below.

The {\it off--zenith} pointing {\bf has notably higher power in the {\it foreground wedge} than in the {\it zenith} pointing}. This is shown later in \S\ref{sec:diffuse} to be due to response of the array to the bright Galactic center and Galactic plane in the westward sky. Consequently, the contamination into the {\it EoR window} is also higher, especially on short antenna spacings. Emission in the {\it zenith} pointing is more centrally concentrated {\bf due to a zenith--centered power pattern and faster suppression of response away from zenith}, whereas the {\it off--zenith} pointing has a power pattern phased eastward of zenith and is also more responsive to emission farther away from the primary lobe relative to that in the {\it zenith} pointing. {\bf The power pattern is discussed in detail in \S\ref{sec:instrument_model}}.

@@

The instrument model consists primarily of the tile power pattern, $A(\hat{\boldsymbol{s}},f)$ (see equation~\ref{eqn:obsvis}). It is modeled as a mutually--coupled 4--by--4 dipole array with the overall power pattern of each individual dipole calculated via finite element electromagnetic simulation \citep{sut14_arXiv}. To speed up simulations, we find that a phased array of isotropic radiators at a height of 0.3~m above an infinite ground plane provides a very good approximation to the full simulation, hence we use the idealized dipoles. We also assume that each individual dipole signal has random delay fluctuations of rms 0.05~ns, a number in line with the known repeatability and stability level of the analog signal chain \citep{bow07b}. Besides having the effect of adding a time--dependent uncertainty in the power pattern, these random delay fluctuations reduce the coherence in the phased addition of dipole signals resulting in deviations from predicted models of the power pattern, most prominently at its nulls. The power patterns so obtained are shown in Figure~\ref{fig:power_pattern} for the pointings chosen in our study.

\begin{figure}[htb]
\centering
% \includegraphics[width=\linewidth]{figures/v1_0/delta_array_powerpattern_0.3m_ground_custom.eps}
% \includegraphics[width=\linewidth]{f4.eps}
\includegraphics[width=\linewidth]{fig4.eps}
\caption{Models of MWA tile power pattern for {\it off--zenith} (top) and {\it zenith} pointings (bottom) at 185~MHz. An MWA tile is modeled as a 4$\times$4 array of isotropic radiators placed 0.3~m above an infinite ground plane. Random fluctuations of rms 0.05~ns have been added to delay corrections during the phased addition of voltages from these isotropic radiators. The logarithmic color scale (shown on the right) is common to both panels. \label{fig:power_pattern}}
\end{figure}

@@

A low resolution version of the diffuse foreground model used is shown in Figure~\ref{fig:DSM}. Contours of the MWA tile power pattern shown in Figure~\ref{fig:power_pattern} are also overlaid. Of notable significance in the {\it off--zenith} pointing is the presence of a portion of the Galactic plane and the bright Galactic center in the westward sky, where the MWA tile power gain is significant ($\gtrsim 12$\%). In the {\it zenith} pointing, the Galactic plane has set and the power pattern in that direction is at least 16 times smaller.

\begin{figure}[htb]
\centering
% \includegraphics[width=\linewidth]{figures/v1_0/dsm.eps}
% \includegraphics[width=\linewidth]{f5.eps}
\includegraphics[width=\linewidth]{fig5.eps}
\caption{Sky brightness temperature of the diffuse foreground model at 185~MHz visible during {\it off--zenith} (top) and {\it zenith} (bottom) pointings. The color scale on the right is logarithmic and is common to both panels. Contours of power pattern shown in Figure~\ref{fig:power_pattern} are overlaid. The contour levels shown are 0.00195, 0.00781, 0.0312, 0.125, and 0.5. The Galactic center and a portion of the Galactic plane are prominently visible during the {\it off--zenith} pointing and the MWA tile power gain is significant ($\gtrsim 12$\%) in that direction. In contrast, emission from the Galactic plane in {\it zenith} pointing is significantly lesser. \label{fig:DSM}}
\end{figure}

@@

\begin{figure}[htb]
\centering
% \includegraphics[width=\linewidth]{figures/v1_0/csm.eps}
% \includegraphics[width=\linewidth]{f6.eps}
\includegraphics[width=\linewidth]{fig6.eps}
\caption{Flux densities of bright point sources at 185~MHz visible during {\it off--zenith} (top) and {\it zenith} (bottom) pointings. The color scale on the right is logarithmic and is common to both panels and corresponds to the color--coded filled circles. Contours of power pattern shown in Figure~\ref{fig:power_pattern} are overlaid. The contour levels shown are 0.001953125, 0.0078125, 0.03125, 0.125, and 0.5. \label{fig:CSM}}
\end{figure}

@@

In addition to intrinsic model uncertainty, delay spectra from simulations and data each have fluctuations due to thermal noise in the delay spectrum with rms $\sim$ 1.4$\times 10^7$~Jy~Hz. We estimate the ratio of delay spectra from data and simulations as $\rho = |V^\textrm{D}_{b\tau}(\boldsymbol{b},\tau)|\,/\,|V^\textrm{S}_{b\tau}(\boldsymbol{b},\tau)|$, where superscripts D and S denote data and simulation, respectively. The median absolute deviation of $\log_{10}\rho$ inside the {\it foreground wedge} for both pointings is $\approx 0.28$. This corresponds to $\sim$~90\% fractional difference between data and modeling on average with either pointing.

We also simulated delay spectra after assigning spectral indices drawn randomly from a {\it gaussian} distribution with a mean of $\langle\alpha_\textrm{sp}\rangle=-0.83$ and a HWHM of $\Delta\alpha_\textrm{sp}=0.35$ to the point sources in our compact foreground model. These simulations typically yielded a median absolute deviation of $\approx 0.29$ for $\log_{10}\rho$ indicating fractional differences of $\sim$~95\%. This demonstrates that a fractional deviation of $\sim$~90\% observed between data and simulations is in line with expectations when the aforementioned uncertainty in foreground models, thermal noise fluctuations in measurements, and uncertainties in antenna power pattern due to random delay fluctuations are taken into account. Currently, we are primarily limited by the unavailability of all--sky foreground models at the observing frequency of 185~MHz.

These uncertainties are presented only to confirm the qualitative agreement already seen between Figures~\ref{fig:fhd_data} and \ref{fig:sim_data}. These estimates are conservative. A full treatment of all uncertainties and deviations from ideal behavior such as frequency dependent errors in tile power pattern \citep{ber14}, calibration \citep{dat10}, data corruption due to interference, anisoplanatic wide--field imaging and ionospheric effects \citep{int09} will bring the simulations much closer in agreement with observations, but is beyond the scope of this paper. {\bf Hereafter}, our primary objective is to explore in detail the foreground signatures embedded in the {\it foreground wedge}.

@@

\item {\it Baseline Orientation}: Since the spatial structure of our foreground model is not expected to be isotropic, we divide our antenna spacings by their orientation, $\theta_\textrm{b}$. We use the {\bf four following bins sweeping 45\arcdeg each}: $-67.5\arcdeg\le \theta_\textrm{b} < -22.5\arcdeg$, $-22.5\arcdeg\le \theta_\textrm{b} < 22.5\arcdeg$, $22.5\arcdeg\le \theta_\textrm{b} < 67.5\arcdeg$, and $67.5\arcdeg\le \theta_\textrm{b} < 112.5\arcdeg$. The bin centers are oriented towards South--East, East, North--East, and North respectively. {\bf While the bin boundaries are arbitrary, they were so chosen to allow bin centers to be aligned along the cardinal and ordinal directions in a half--plane}. Since delays depend on $\theta_\textrm{b}$, binning by $\theta_\textrm{b}$ allows us to match {\bf features of the delay power spectra} to different sky directions.

\item {\it Tile Pointing and Power Pattern}: Since the MWA tiles are steered electronically, the tile power pattern tile changes in any observing mode that tracks the {\bf pointing center}. In our study, we take into account the effect of the changing tile power pattern on the {\bf delay power spectra} (Figure~\ref{fig:power_pattern}). The LST of our data ranges from $\sim$~21 hours through $\sim$~2 hours, and for our analysis we chose the {\it off--zenith} and {\it zenith} pointings shown in Figure~\ref{fig:pointings}. % We will show the secondary lobe structures in the tile power pattern are tied to the foreground signatures in Fourier space.

@@

In subsequent sections, we provide a detailed explanation of our results as a combination of factors noted above. Note that numerous features may overlap at different degrees of significance depending on combinations of parameters. We assign the features to their predominant causes. Secondly, we have used noiseless cases to clearly illustrate the observed foreground signatures. With the addition of noise in the visibilities, some of the weaker features may not be as prominently visible. Since the foreground signatures are far too numerous {\bf and subject to a multitude of parameters like} baseline length and orientation, power pattern, patch of sky under observation, and instrumental configuration, we highlight only some examples of the most notable features in the delay spectra of foreground emission.

@@

The most prominent signature seen in the {\it off--zenith} pointing (top panel of Figure~\ref{fig:noiseless-dsm-delay-spectrum}) is due to the bright Galactic center situated on the western horizon co--located with one of the bright secondary lobes of the power pattern. It appears as a bright branch near the negative delay horizon delay limit. This feature is strongest at short antenna spacings and fades with increasing antenna spacing. The {\bf bright} signature is absent in the {\it zenith} pointing (bottom panel, Figure~\ref{fig:noiseless-dsm-delay-spectrum}) because the Galactic center is below the horizon.

@@

The amplitude response of an interferometer to a point source is, to first order, flat with respect to baseline length. Since the primary lobe of the power pattern in the {\it off--zenith} pointing is centered eastward of zenith, the bulk of the compact foreground emission is seen in a branch with $\tau>0$ corresponding to the position of the primary lobe of the power pattern. In the {\it zenith} pointing, compact emission from the same patch of sky is seen as a bright horizontal arm at $\tau=0$ since the primary lobe of the power pattern is centered at zenith.
